---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tblrelations

> Could these data frames be a database?

[![Build Status](https://travis-ci.org/jalsalam/tblrelations.svg?branch=master)](https://travis-ci.org/jalsalam/tblrelations)

## Installation

You can install the development version of tblrelations from Github with:

``` r
devtools::install_github("jalsalam/tblrelations")
```

## Example

When working with data frames in R in a relational data way, it is useful to identify primary and foreign keys:

```{r flights}
library(tblrelations)
suppressPackageStartupMessages(library(dplyr))
library(nycflights13)

airports
```

```{r cars}
pk_ish(airports, by = "faa")
```

```{r}
flights
```

```{r}
pk_ish(flights, by = c("year", "month", "day", "flight"))
pk_ish(flights, by = c("year", "month", "day", "tailnum"))
pk_ish(flights, by = c("year", "month", "day", "hour", "flight"))
pk_ish(flights, by = c("year", "month", "day", "hour", "carrier", "flight"))
pk_ish(flights, by = c("year", "month", "day", "carrier", "flight"))
pk_ish(flights, by = c("year", "month", "day", "origin", "carrier", "flight"))

# pk_ish(flights, .keys = vars(year, month, day, carrier, flight)) # currently errors
```

See: [r4ds relational data](http://r4ds.had.co.nz/relational-data.html))


```{r}
fk_ish(flights, airlines, by = "carrier")
#fk_ish(flights, airports, by = c("origin" = "faa")) # currently errors
```


```{r band_members}

pk_ish(dplyr::band_members, by = "name")
# pk_ish(dplyr::band_members, by = vars(name)) # currently errors

fk_ish(band_members, band_instruments, by = "name")
# fk_ish(band_members, band_instruments2, by = c("name" = "artist")) # currently errors
```

Note: it should be possible to use rename_at like options:

```{r}
#rename_at(band_instruments2, vars(name = "artist")) #error b/c you must rename with a function
rename_at(band_instruments2, vars("name" = "artist", "instrument" = "plays"), funs(names))
rename_at(band_instruments2, vars(name = "artist", instrument = "plays"), funs(names))
rename_at(band_instruments2, c("name" = "artist", "instrument" = "plays"), funs(names))
```





Have you ever been surprised that a left_join resulted in an unexpected cross-join or NA rows?

``` r
pk_ish(dfy, by = "id")
# TRUE if dfy$pk is unique
# FALSE if dfy$pk is not unique

fk_ish(dfx, dfy, by = "id")
# TRUE if all combination of "id" in dfx are in dfy and dfy$id is unique.

left_join_fk(dfx, dfy, by = "id")
# if there is a foreign-key-ish relationship of dfx -> dfy using "id", then it returns the joined table as normal, otherwise it errors.

```

## Development goals

1. Make sure that the same semantics of the `by` argument can be used with `left_join_fk()` as can be used with the normal dplyr join verbs. Using rename_at like semantics, it should eb possible to accept either a character vector or a `vars` specification.

2. Add diagnostics and good error messages that let you know if there are missing or duplicate values on your dimension table. Maybe this should be under an option `verbose = TRUE` or something.

3. Add an assertion version of some o fthese checks -- e.g., a good explanatory error message for use with testthat.

4. Add right_join_fk. Are there other join relationships that make sense for these kinds of checks?


## Much more speculative. Really need some DB research on what else might be helpful:

- Maybe s3 object for which you can more permanently label pk?

- Maybe express domains, codomains, onto, covers, or other such function/set theory types of concepts as tests on relations.

- Maybe coerce a table to a function. E.g. :

```r
join_dfy <- as_tbl_function(dfy, by = "id")

dfx %>% join_dfy(by = "id")
```

But there are plenty of questions about what else would be worthwhile.

